---
title: "ofav_mcav_sctld_count_import"
output: html_document
---

I always load all my packages as the first chunk in every rmarkdown I make. Its nicer than having them splayed around and also if you are knitting to a html then it means you will have no package dependency problems anywhere in the document. 

```{r Package Loading, include = F}
library(tximport)
library(DESeq2)
library(tidyverse)
library(tximport)
library(ComplexHeatmap)
library(DEGreport)
library(limma)
library(edgeR)
library(pheatmap)
```


## Importing Salmon Quant Files

The quant files are all on the box 

**PLEASE DO NOT UPLOAD THESE TO ANY PUBLIC PLACES AS THIS IS THE RAW DATA**

We are going to download the quant files so you can make the count tables yourself. To do that we are going to have a folder on the desktop as this means you will not have to chnage any paths.You just need to make sure everything is saved in the correct folders 

```{bash}
cd ~/Desktop
mkdir SCTLD_analysis
```

```{bash}
cd ~/Desktop/SCTLD_analysis
mkdir scripts
```

Once you have done the above go to the box and download the  
+ ofav_salmon_quant/
+ mcav_salmon_quant/
+ metadata.csv

These should be saved in the SCTLD_analysis folder so you should have a directory tree as such 

Desktop/
      |
      ofav_salmon_quant/
      mcav_salmon_quant/
      scripts/
      metadata.csv
      

```{r Making Salmon Vectors Names ofav, include=FALSE}
#Loading quantification data output from the slippery Salmon
ofavall <- list.files(path = "~/Desktop/SCTLD_analysis/ofav_salmon_quant//", full.names = F, pattern = "\\_salmon$")

FILESofav <- file.path(ofavall, "quant.sf")

names(FILESofav) <- ofavall
head(FILESofav)
all(file.exists(FILESofav))
```

```{r Making Salmon Vectors Names mcav, include=FALSE}
#Loading quantification data output from the slippery Salmon
setwd("~/Desktop/SCTLD_analysis/mcav_salmon_quant/")
mcavall <- list.files(path = "~/Desktop/SCTLD_analysis/mcav_salmon_quant//", full.names = F, pattern = "\\_salmon$")

FILESmcav <- file.path(mcavall, "quant.sf")

names(FILESmcav) <- mcavall
head(FILESmcav)
all(file.exists(FILESmcav))
```

```{r Reading in Salmon Files ofav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/SCTLD_analysis/ofav_salmon_quant//")
txi.salmon.count_ofav <- tximport(FILESofav, type = "salmon", txOut = TRUE )
```

```{r Reading in Salmon Files mcav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/SCTLD_analysis/mcav_salmon_quant//")
txi.salmon.count_mcav <- tximport(FILESmcav, type = "salmon", txOut = TRUE )
```

We have read these in so lets just save all our count, abundance and length files as well

```{bash}
cd ~/Desktop/SCTLD_analysis/
mkdir count_folder
```

```{r writing csv for all info mcav, include = F}
write.csv(txi.salmon.count_mcav$counts, file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_counts.csv")
write.csv(txi.salmon.count_mcav$abundance , file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_abundance.csv")
write.csv(txi.salmon.count_mcav$length, file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_length.csv")
```


```{r writing csv for all info ofav, include = F}
write.csv(txi.salmon.count_ofav$counts, file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_counts.csv")
write.csv(txi.salmon.count_ofav$abundance , file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_abundance.csv")
write.csv(txi.salmon.count_ofav$length, file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_length.csv")
```

Now we are going to read in the metadata file and fix these for both species

```{r}
meta <- read.csv(file = "~/Desktop/SCTLD_analysis/metadata.csv")
str(meta)
```

Here, I filter the files I imported above and then filter them with the metadata file. This is needed as libraries failed and we only want the data for the samples that worked.  

```{r filter mcav}
meta %>% filter(salmon_label %in% mcavall) -> mcav_meta

## Check our salmon and meta lengths the same
length(mcavall)
nrow(mcav_meta)
```

```{r filter ofav}
meta %>% filter(salmon_label %in% ofavall) -> ofav_meta
length(ofavall)
nrow(ofav_meta)
```

## Making DeSeq2 Object 

```{r Making DeSeq object, include=FALSE}
#making the dds model to use in deseq2
dds_mcav = DESeqDataSetFromTximport(txi.salmon.count_mcav , mcav_meta, ~ Correct_treatment)
dds_ofav = DESeqDataSetFromTximport(txi.salmon.count_ofav , ofav_meta, ~ Correct_treatment)
```













