---
title: "ofav_mcav_sctld_count_import"
output: html_document
---

I always load all my packages as the first chunk in every rmarkdown I make. Its nicer than having them splayed around and also if you are knitting to a html then it means you will have no package dependency problems anywhere in the document. 

```{r Package Loading, include = F}
library(tximport)
library(DESeq2)
library(tidyverse)
library(tximport)
library(ComplexHeatmap)
#library(DEGreport)
library(limma)
library(edgeR)
library(pheatmap)
```


## Importing Salmon Quant Files

The quant files are all on the box 

**PLEASE DO NOT UPLOAD THESE TO ANY PUBLIC PLACES AS THIS IS THE RAW DATA**

We are going to download the quant files so you can make the count tables yourself. To do that we are going to have a folder on the desktop as this means you will not have to chnage any paths.You just need to make sure everything is saved in the correct folders 

```{bash}
cd ~/Desktop/sctld_jamboree
mkdir SCTLD_analysis
```

```{bash}
cd ~/Desktop/sctld_jamboree/SCTLD_analysis
mkdir scripts
```

Once you have done the above go to the box and download the  
+ ofav_salmon_quant/
+ mcav_salmon_quant/
+ metadata.csv

These should be saved in the SCTLD_analysis folder so you should have a directory tree as such 

Desktop/
      |
      ofav_salmon_quant/
      mcav_salmon_quant/
      scripts/
      metadata.csv
      

```{r Making Salmon Vectors Names ofav, include=FALSE}
#Loading quantification data output from the slippery Salmon
ofavall <- list.files(path = "~/Desktop/ofav_salmon_quant//", full.names = F, pattern = "\\_salmon$")

FILESofav <- file.path(ofavall, "quant.sf")

names(FILESofav) <- ofavall
head(FILESofav)
all(file.exists(FILESofav))
```

```{r Making Salmon Vectors Names mcav, include=FALSE}
#Loading quantification data output from the slippery Salmon
setwd("~/Desktop/mcav_salmon_quant/")
mcavall <- list.files(path = "~/Desktop/mcav_salmon_quant//", full.names = F, pattern = "\\_salmon$")

FILESmcav <- file.path(mcavall, "quant.sf")

names(FILESmcav) <- mcavall
head(FILESmcav)
all(file.exists(FILESmcav))
```

```{r Reading in Salmon Files ofav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/ofav_salmon_quant//")
txi.salmon.count_ofav <- tximport(FILESofav, type = "salmon", txOut = TRUE )
```

```{r Reading in Salmon Files mcav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/mcav_salmon_quant//")
txi.salmon.count_mcav <- tximport(FILESmcav, type = "salmon", txOut = TRUE )
```

We have read these in so lets just save all our count, abundance and length files as well

```{bash}
cd ~/Desktop
mkdir SCTLD_analysis
cd ~/Desktop/SCTLD_analysis/
mkdir count_folder
```

```{r writing csv for all info mcav, include = F}
write.csv(txi.salmon.count_mcav$counts, file = "~/Desktop/SCTLD_analysis/count_folder/mcav_counts.csv")
write.csv(txi.salmon.count_mcav$abundance , file = "~/Desktop/SCTLD_analysis/count_folder/mcav_abundance.csv")
write.csv(txi.salmon.count_mcav$length, file = "~/Desktop/SCTLD_analysis/count_folder/mcav_length.csv")
```


```{r writing csv for all info ofav, include = F}
write.csv(txi.salmon.count_ofav$counts, file = "~/Desktop/SCTLD_analysis/count_folder/ofav_counts.csv")
write.csv(txi.salmon.count_ofav$abundance , file = "~/Desktop/SCTLD_analysis/count_folder/ofav_abundance.csv")
write.csv(txi.salmon.count_ofav$length, file = "~/Desktop/SCTLD_analysis/count_folder/ofav_length.csv")
```

Now we are going to read in the metadata file and fix these for both species

```{r}
meta <- read.csv(file = "~/Desktop/SCTLD_analysis/metadata.csv")
str(meta)
```

Here, I filter the files I imported above and then filter them with the metadata file. This is needed as libraries failed and we only want the data for the samples that worked.  

```{r filter mcav}
meta %>% filter(salmon_label %in% mcavall) -> mcav_meta

## Check our salmon and meta lengths the same
length(mcavall)
nrow(mcav_meta)
```

```{r filter ofav}
meta %>% filter(salmon_label %in% ofavall) -> ofav_meta
length(ofavall)
nrow(ofav_meta)
```

## Making DeSeq2 Object 

```{r Making DeSeq object, include=FALSE}
#making the dds model to use in deseq2
dds_mcav = DESeqDataSetFromTximport(txi.salmon.count_mcav , mcav_meta, ~ Correct_treatment)
dds_ofav = DESeqDataSetFromTximport(txi.salmon.count_ofav , ofav_meta, ~ Correct_treatment)
```

```{r}
head(dds_mcav)
head(dds_ofav)
```
```{r}
# filter out low counts
keep_mcav <- rowSums(counts(dds_mcav)) >=10
keep_ofav <- rowSums(counts(dds_ofav)) >=10

dds_mcav_rmlow <- dds_mcav[keep_mcav,]
dds_ofav_rmlow <- dds_ofav[keep_ofav,]
```
How many non-zero genes?
```{r}
GeneCountsMcav <- counts(dds_mcav_rmlow)
idx.nz1 <- apply(GeneCountsMcav, 1, function(x) { all(x > 0)})
sum(idx.nz1)

GeneCountsOfav <- counts(dds_ofav_rmlow)
idx.nz2 <- apply(GeneCountsOfav, 1, function(x) { all(x > 0)})
sum(idx.nz2)
```
REF: http://www.nathalievialaneix.eu/doc/html/TP1_normalization.html 
```{r}
library(ggplot2)
library(gridExtra)
library(reshape2)
#library(mixOmics)
library(RColorBrewer)
library(DESeq2)
library(edgeR)
library(VennDiagram)
library(devtools)
```
Make a data frame of the counts for normalization with edgeR
```{r}
ofav_counts <- as.data.frame(txi.salmon.count_ofav$counts)
mcav_counts <- as.data.frame(txi.salmon.count_mcav$counts)

ofav_lengths <- as.data.frame(txi.salmon.count_ofav$length)
mcsv_lengths <- as.data.frame(txi.salmon.count_mcav$length)
```
EdgeR used TMM normalization ()
```{r}
# filter out reads with count of 0
raw_counts_ofav <- ofav_counts[rowSums(ofav_counts) > 0,]
raw_counts_mcav <- mcav_counts[rowSums(mcav_counts) > 0, ]


# calucluate log transformed coutns
pseudo_counts_ofav <- log2(raw_counts_ofav + 1)
head(pseudo_counts_ofav)

pseudo_counts_mcav <- log2(raw_counts_mcav + 1)
head(pseudo_counts_mcav)

#ofav
df_raw_ofav <- melt(as.matrix(pseudo_counts_ofav), id = rownames(raw_counts_ofav))
names(df_raw_ofav)[1:2]<- c("id", "sample")
df_raw_ofav$method <- rep("Raw counts", nrow(df_raw_ofav))  

#mcav
df_raw_mcav <- melt(as.matrix(pseudo_counts_mcav), id = rownames(raw_counts_mcav))
names(df_raw_mcav)[1:2]<- c("id", "sample")
df_raw_mcav$method <- rep("Raw counts", nrow(df_raw_mcav))  
```
Visualize distribution of counts
```{r}
#ofav
df <- data.frame(rcounts = raw_counts_ofav[ ,1], prcounts = pseudo_counts_ofav[ ,1])

p <- ggplot(data=df, aes(x = rcounts, y = ..density..))
p <- p + geom_histogram(fill = "lightblue")
p <- p + theme_bw()
p <- p + ggtitle(paste0("count distribution ofav"))
p <- p + xlab("counts")

p2 <- ggplot(data=df, aes(x = prcounts, y = ..density..))
p2 <- p2 + geom_histogram(fill = "lightblue")
p2 <- p2 + theme_bw()
p2 <- p2 + ggtitle(paste0("count distribution ofav log2"))
p2 <- p2 + xlab(expression(log[2](counts + 1)))

#mcav
df2 <- data.frame(rcounts = raw_counts_mcav[ ,1], prcounts = pseudo_counts_mcav[ ,1])

p3 <- ggplot(data=df2, aes(x = rcounts, y = ..density..))
p3 <- p + geom_histogram(fill = "lightblue")
p3 <- p + theme_bw()
p3 <- p + ggtitle(paste0("count distribution mcav"))
p3 <- p + xlab("counts")

p4 <- ggplot(data=df2, aes(x = prcounts, y = ..density..))
p4 <- p2 + geom_histogram(fill = "lightblue")
p4 <- p2 + theme_bw()
p4 <- p2 + ggtitle(paste0("count distribution mcav log2"))
p4 <- p2 + xlab(expression(log[2](counts + 1)))

grid.arrange(p, p2, p3, p4, ncol = 2, nrow=2)
```
Look at variance and means because RNA data is overdispersed (i.e. variance in read counts will be higher than read counts)
```{r}
#ofav 
df <- data.frame(mean = apply(raw_counts_ofav[ ,ofav_meta$Correct_treatment == "control"], 1, mean),
                 var = apply(raw_counts_ofav[ ,ofav_meta$Correct_treatment == "control"], 1, var))
p <- ggplot(data=df, aes(x = mean, y = var))
p <- p + geom_point(colour = "orange")
p <- p + theme_bw()
p <- p + geom_abline(aes(intercept=0, slope=1))
p <- p + ggtitle("Variance versus mean in counts ofav") + ylab("variance")
print(p)

#mcav
df2 <- data.frame(mean = apply(raw_counts_mcav[ ,mcav_meta$Correct_treatment == "control"], 1, mean),
                 var = apply(raw_counts_mcav[ ,mcav_meta$Correct_treatment == "control"], 1, var))
p2 <- ggplot(data=df2, aes(x = mean, y = var))
p2 <- p2 + geom_point(colour = "orange")
p2 <- p2 + theme_bw()
p2 <- p2 + geom_abline(aes(intercept=0, slope=1))
p2 <- p2 + ggtitle("Variance versus mean in counts mcav") + ylab("variance")
print(p2)

#ofav 
df <- data.frame(mean = apply(raw_counts_ofav[ ,ofav_meta$Correct_treatment == "control"], 1, mean),
                 var = apply(raw_counts_ofav[ ,ofav_meta$Correct_treatment == "control"], 1, var))
df <- df[df$mean <= 5000, ]
p <- ggplot(data=df, aes(x = mean, y = var))
p <- p + geom_point(colour = "orange")
p <- p + theme_bw()
p <- p + geom_abline(aes(intercept=0, slope=1))
p <- p + ggtitle("Variance versus mean in counts ofav") + ylab("variance")
print(p)

#mcav
df2 <- data.frame(mean = apply(raw_counts_mcav[ ,mcav_meta$Correct_treatment == "control"], 1, mean),
                 var = apply(raw_counts_mcav[ ,mcav_meta$Correct_treatment == "control"], 1, var))
df2 <- df2[df2$mean <= 5000, ]
p2 <- ggplot(data=df2, aes(x = mean, y = var))
p2 <- p2 + geom_point(colour = "orange")
p2 <- p2 + theme_bw()
p2 <- p2 + geom_abline(aes(intercept=0, slope=1))
p2 <- p2 + ggtitle("Variance versus mean in counts mcav") + ylab("variance")
print(p2)
```
Do a basic PCA to determine sources of variance
```{r}
library(factoextra)
#ofav
res.pca <- prcomp(t(pseudo_counts_ofav), scale = TRUE)
fviz_eig(res.pca)
groups <- as.factor(ofav_meta$Correct_treatment)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)

#mcav
res.pca2 <- prcomp(t(pseudo_counts_mcav), scale = TRUE)
fviz_eig(res.pca2)
groups <- as.factor(mcav_meta$Correct_treatment)
fviz_pca_ind(res.pca2, col.ind=groups, label=FALSE)
```
## Normalization ##
```{r}
# Use DESeq - corrects for library size NOT gene specific
#ofav
sizefact <- estimateSizeFactorsForMatrix(raw_counts_ofav)
sizefact_dd <- estimateSizeFactors(dds_ofav)
deseq_normcount <- counts(sizefact_dd, normalized = TRUE)
test_normcount <- sweep(raw_counts_ofav, 2, sizefact, FUN = "/")
sum(test_normcount != deseq_normcount)

pseudo_deseq <- log2(deseq_normcount + 1)
df_deseq <- melt(pseudo_deseq, id = rownames(raw_counts_ofav))
names(df_deseq)[1:2]<- c("id", "sample")
df_deseq$method <- rep("DESeq (RLE)", nrow(df_deseq)) 

#mcav
sizefact2 <- estimateSizeFactorsForMatrix(raw_counts_mcav)
sizefact_dd2 <- estimateSizeFactors(dds_mcav)
deseq_normcount2 <- counts(sizefact_dd2, normalized = TRUE)
test_normcount2 <- sweep(raw_counts_mcav, 2, sizefact2, FUN = "/")
sum(test_normcount2 != deseq_normcount2)

pseudo_deseq2 <- log2(deseq_normcount2 + 1)
df_deseq2 <- melt(pseudo_deseq2, id = rownames(raw_counts_mcav))
names(df_deseq2)[1:2]<- c("id", "sample")
df_deseq2$method <- rep("DESeq (RLE)", nrow(df_deseq2)) 
```

```{r}
# EdgeR
#ofav
dge2 <- DGEList(raw_counts_ofav)
dge2

dge2$samples
dge2_keep <- filterByExpr(dge2, group=ofav_meta$Correct_treatment, min.count=34)
summary(dge2_keep)
dge2 <- dge2[dge2_keep, , keep.lib.sizes=FALSE]

#mcav
dge3 <- DGEList(raw_counts_mcav)
dge3

dge3$samples
dge3_keep <- filterByExpr(dge3, group=mcav_meta$Correct_treatment, min.count=15)
summary(dge3_keep)
dge3 <- dge3[dge3_keep, , keep.lib.sizes=FALSE]
```
Normalization method: total count
```{r}
#ofav
pseudo_TC <- log2(cpm(dge2) + 1)

df_TC <- melt(pseudo_TC, id = rownames(raw_counts_ofav))
names(df_TC)[1:2] <- c ("id", "sample")
df_TC$method <- rep("TC", nrow(df_TC))

# mcav
pseudo_TC2 <- log2(cpm(dge3) + 1)

df_TC2 <- melt(pseudo_TC2, id = rownames(raw_counts_mcav))
names(df_TC2)[1:2] <- c ("id", "sample")
df_TC2$method <- rep("TC", nrow(df_TC2))
```
Normalization method: reads per thousand counts (RPKM)
```{r}
#ofav
ofav_lengths_use <- ofav_lengths[rowSums(ofav_counts) > 0,]
pseudo_RPKM <- log2(rpkm(dge2, gene.length = ofav_lengths_use) + 1)

df_RPKM <- melt(as.matrix(pseudo_RPKM), id.vars = rownames(raw_counts_ofav))
names(df_RPKM)[1:2] <- c ("id", "sample")
df_RPKM$method <- rep("RPKM", nrow(df_RPKM))

#mcav
mcav_lengths_use <- mcsv_lengths[rowSums(mcav_counts) > 0,]
pseudo_RPKM2 <- log2(rpkm(dge3, gene.length = mcav_lengths_use) + 1)

df_RPKM2 <- melt(as.matrix(pseudo_RPKM2), id = rownames(raw_counts_mcav))
names(df_RPKM2)[1:2] <- c ("id", "sample")
df_RPKM2$method <- rep("RPKM", nrow(df_RPKM2))
```
Normaliztion method: Upper Quartile, not working, getting inf or nan for norm factor
```{r}
#ofav
dge2 <- calcNormFactors(dge2, method = "upperquartile")
dge2$samplesdge3_keep <- filterByExpr(dge3, group=mcav_meta$Correct_treatment, min.count=15)
dge3 <- dge3[dge3_keep, , keep.lib.sizes=FALSE]

test_normcount <- sweep(dge2$counts, 2,
                        dge2$samples$lib.size*dge2$samples$norm.factors / 10^6,
                        FUN="/")
range(as.vector(test_normcount - cpm(dge2)))

pseudo_UQ <- log2(cpm(dge2) + 1)

df_UQ <- melt(as.matrix(pseudo_UQ), id = rownames(raw_counts_ofav))
names(df_UQ)[1:2] <- c ("id", "sample")
df_UQ$method <- rep("UQ", nrow(df_UQ))

#mcav
dge3 <- calcNormFactors(dge3, method = "upperquartile")
dge3$samples
dge3_keep <- filterByExpr(dge3, group=mcav_meta$Correct_treatment)
dge3 <- dge3[dge3_keep, , keep.lib.sizes=FALSE]
test_normcount2 <- sweep(dge3$counts, 2,
                        dge3$samples$lib.size*dge3$samples$norm.factors / 10^6,
                        FUN="/")
range(as.vector(test_normcount2 - cpm(dge3)))

pseudo_UQ2 <- log2(cpm(dge3) + 1)

df_UQ2 <- melt(as.matrix(pseudo_UQ2), id = rownames(raw_counts_mcav))
names(df_UQ2)[1:2] <- c ("id", "sample")
df_UQ2$method <- rep("UQ", nrow(df_UQ2))
```
Normalization method: TMM
```{r}
#ofav
dge2 <- calcNormFactors(dge2, method = "TMM")
dge2$samples
dge2_keep <- filterByExpr(dge2, group=ofav_meta$Correct_treatment, min.count=34)
dge2 <- dge2[dge2_keep, , keep.lib.sizes=FALSE]
pseudo_TMM <- log2(cpm(dge2) + 1)

df_TMM <- melt(as.matrix(pseudo_TMM), id = rownames(raw_counts_ofav))
names(df_TMM)[1:2] <- c ("id", "sample")
df_TMM$method <- rep("TMM", nrow(df_TMM))

#mcav
dge3 <- calcNormFactors(dge3, method = "TMM")
dge3$samples
dge3_keep <- filterByExpr(dge3, group=mcav_meta$Correct_treatment, min.count=15)
dge3 <- dge3[dge3_keep, , keep.lib.sizes=FALSE]
pseudo_TMM2 <- log2(cpm(dge3) + 1)

df_TMM2 <- melt(as.matrix(pseudo_TMM2), id = rownames(raw_counts_mcav))
names(df_TMM2)[1:2] <- c ("id", "sample")
df_TMM2$method <- rep("TMM", nrow(df_TMM2))
```
Compare all normalization methods 
```{r, fig.width=10}
# Plots for ofav
df_allnorm <- rbind(df_raw_ofav, df_deseq, df_TC, df_RPKM, df_TMM)
df_allnorm$method <- factor(df_allnorm$method,
                            levels = c("Raw counts", "DESeq (RLE)", "TC", 
                                       "RPKM", "TMM"))

p <- ggplot(data=df_allnorm, aes(x=sample, y=value, fill=method))
p <- p + geom_boxplot()  
p <- p + theme_bw()
p <- p + ggtitle("Boxplots of normalized pseudo counts\n
for all samples by normalization methods - ofav")
p <- p + facet_grid(. ~ method) 
p <- p + ylab(expression(log[2] ~ (normalized ~ count + 1))) + xlab("")
p <- p + theme(title = element_text(size=10), axis.text.x = element_blank(), 
               axis.ticks.x = element_blank())
print(p)
```

```{r, fig.width=10}
# Plots for mcav
df_allnorm2 <- rbind(df_raw_mcav, df_deseq2, df_TC2, df_RPKM2, df_TMM2)
df_allnorm2$method <- factor(df_allnorm2$method,
                            levels = c("Raw counts", "DESeq (RLE)", "TC", 
                                       "RPKM", "TMM"))

p2 <- ggplot(data=df_allnorm2, aes(x=sample, y=value, fill=method))
p2 <- p2 + geom_boxplot()  
p2 <- p2 + theme_bw()
p2 <- p2 + ggtitle("Boxplots of normalized pseudo counts\n
for all samples by normalization methods - mcav")
p2 <- p2 + facet_grid(. ~ method) 
p2 <- p2 + ylab(expression(log[2] ~ (normalized ~ count + 1))) + xlab("")
p2 <- p2 + theme(title = element_text(size=10), axis.text.x = element_blank(), 
               axis.ticks.x = element_blank())
print(p2)
```
```{r, fig.width=10}
#ofav
p <- ggplot(data=df_allnorm, aes(x=value, colour=sample))
p <- p + geom_density()  
p <- p + theme_bw()
p <- p + ggtitle("Density of normalized pseudo counts\n
for all samples by normalization methods - ofav")
p <- p + facet_grid(. ~ method) 
p <- p + ylab(expression(log[2] ~ (normalized ~ count + 1))) + xlab("")
p <- p + theme(title = element_text(size=10), axis.text.x = element_blank(), 
               axis.ticks.x = element_blank())
print(p)
```

```{r, fig.width=10}
#mcav
p2 <- ggplot(data=df_allnorm2, aes(x=value, colour=sample))
p2 <- p2 + geom_density()  
p2 <- p2 + theme_bw()
p2 <- p2 + ggtitle("Density of normalized pseudo counts\n
for all samples by normalization methods - mcav")
p2 <- p2 + facet_grid(. ~ method) 
p2 <- p2 + ylab(expression(log[2] ~ (normalized ~ count + 1))) + xlab("")
p2 <- p2 + theme(title = element_text(size=10), axis.text.x = element_blank(), 
               axis.ticks.x = element_blank())
print(p2)
```
Repeat PCA with normalized TMM data
```{r}
#ofav
res.pca <- prcomp(t(pseudo_RPKM), scale = TRUE)
fviz_eig(res.pca)
groups <- as.factor(ofav_meta$Correct_treatment)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)

res.pca <- prcomp(t(pseudo_TMM), scale = TRUE)
fviz_eig(res.pca)
groups <- as.factor(ofav_meta$Correct_treatment)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)


groups <- as.factor(ofav_meta$Project)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)


groups <- as.factor(ofav_meta$Exp_frag_genet)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)
```

```{r}
#mcav
res.pca2 <- prcomp(t(pseudo_RPKM2), scale = TRUE)
fviz_eig(res.pca2)
groups <- as.factor(mcav_meta$Correct_treatment)
fviz_pca_ind(res.pca2, col.ind=groups, label=FALSE)

#pc4 is ok
res.pca2 <- prcomp(t(pseudo_TMM2), scale = TRUE)
fviz_eig(res.pca2)
groups <- as.factor(mcav_meta$Correct_treatment)
fviz_pca_ind(res.pca2, col.ind=groups, label=FALSE)

res.pca2 <- prcomp(t(pseudo_TMM2), scale = TRUE)
fviz_eig(res.pca2)
groups <- as.factor(mcav_meta$Project)
fviz_pca_ind(res.pca2, col.ind=groups, label=FALSE)

res.pca <- prcomp(t(pseudo_TMM2), scale = TRUE)
fviz_eig(res.pca)
groups <- as.factor(mcav_meta$Exp_frag_genet)
fviz_pca_ind(res.pca, col.ind=groups, label=FALSE)
```
Try PCA with DESeq2 -- takes top ~500 rows
```{r}
vsd <- vst(dds_ofav, blind=FALSE)
plotPCA(vsd, intgroup="Correct_treatment")

vsd2 <- vst(dds_mcav, blind=FALSE)
plotPCA(vsd2, intgroup="Correct_treatment")
```


