---
title: "ofav_mcav_sctld_count_import"
output: html_document
---

I always load all my packages as the first chunk in every rmarkdown I make. Its nicer than having them splayed around and also if you are knitting to a html then it means you will have no package dependency problems anywhere in the document. 

```{r Package Loading, include = F}
library(tximport)
library(DESeq2)
library(tidyverse)
library(tximport)
library(ComplexHeatmap)
library(DEGreport)
library(limma)
library(edgeR)
library(pheatmap)
```


## Importing Salmon Quant Files

The quant files are all on the box 

**PLEASE DO NOT UPLOAD THESE TO ANY PUBLIC PLACES AS THIS IS THE RAW DATA**

You also do not need to do this as I have generated all the files you will need for analysis. This code is just for your reference on how I do it. 

```{r Making Salmon Vectors Names ofav, include=FALSE}
#Loading quantification data output from the slippery Salmon
setwd("~/Desktop/SCTLD_tagseq/ofav_quant/")
ofavall <- list.files(path = "~/Desktop/SCTLD_tagseq/ofav_quant///", full.names = F, pattern = "\\_salmon$")

FILESofav <- file.path(ofavall, "quant.sf")

names(FILESofav) <- ofavall
head(FILESofav)
all(file.exists(FILESofav))
```

```{r Making Salmon Vectors Names mcav, include=FALSE}
#Loading quantification data output from the slippery Salmon
setwd("~/Desktop/SCTLD_tagseq/mcav_quant/")
mcavall <- list.files(path = "~/Desktop/SCTLD_tagseq/mcav_quant//", full.names = F, pattern = "\\_salmon$")

FILESmcav <- file.path(mcavall, "quant.sf")

names(FILESmcav) <- mcavall
head(FILESmcav)
all(file.exists(FILESmcav))
```

```{r Reading in Salmon Files ofav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/SCTLD_tagseq/ofav_quant//")
txi.salmon.count_ofav <- tximport(FILESofav, type = "salmon", txOut = TRUE )
```

```{r Reading in Salmon Files mcav, include=F}
#Importing the count data using salmon quant.sf files and the text to gene file
setwd("~/Desktop/SCTLD_tagseq/mcav_quant//")
txi.salmon.count_mcav <- tximport(FILESmcav, type = "salmon", txOut = TRUE )
```

```{r writing csv for all info mcav, include = F}
write.csv(txi.salmon.count_mcav$counts, file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_counts.csv")
write.csv(txi.salmon.count_mcav$abundance , file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_abundance.csv")
write.csv(txi.salmon.count_mcav$length, file = "~/Desktop/SCTLD_tagseq/count_folder/mcav_length.csv")
```


```{r writing csv for all info ofav, include = F}
write.csv(txi.salmon.count_ofav$counts, file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_counts.csv")
write.csv(txi.salmon.count_ofav$abundance , file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_abundance.csv")
write.csv(txi.salmon.count_ofav$length, file = "~/Desktop/SCTLD_tagseq/count_folder/ofav_length.csv")
```

















